---
title: "Dplyr tricks"
author: "Suzan Baert"
output:
  xaringan::moon_reader:
    css: ["default", "rladies-fonts", "rladies"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: ""
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

library(dplyr)
```


### Tip 1: Are you often selecting the same columns over and over again?

You can make a vector of pre-identified columns once, and then refer to them using `one_of()` or even shorter with `!!`.

```{r}
library(dplyr)

cols <- c("mpg", "cyl", "gear")

mtcars %>% 
  select(!!cols)

```


---

### Tip 2: Reordering your columns

If you just want to bring one or more columns to the front, you can use `everything()` to add all the remaining columns.

```{r}
library(dplyr)

iris %>% 
  select(Species, everything())
```



---

### Tip 3: Renaming your variables in one go

One command to get them all in lower case. And one more to replace those dots with underscores...


```{r}
library(dplyr)
library(stringr)

iris %>% 
  rename_all(tolower) %>% 
  rename_all(~str_replace_all(., "\\.", "_"))
```


---

