---
title: "Dplyr tricks"
author: "Suzan Baert"
output:
  xaringan::moon_reader:
    css: ["default", "rladies-fonts", "rladies"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: ""
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, message = FALSE, warning = FALSE)

library(dplyr)
```


### Tip 1: Are you often selecting the same columns over and over again?

You can make a vector of pre-identified columns once, and then refer to them using `one_of()` or even shorter with `!!`.

```{r}
library(dplyr)

cols <- c("mpg", "cyl", "gear")

mtcars %>% 
  select(!!cols)

```


---

### Tip 2: Select columns via regex

If you have matching patterns you can use `starts_with()`, `contains()` or `ends_with()`. But what if your pattern isn't thàt exact? Simple: enter regex into `matches()`

```{r}
library(dplyr)

iris %>% 
  select(matches("S.+th"))

```


---

### Tip 3: Reordering your columns

If you just want to bring one or more columns to the front, you can use `everything()` to add all the remaining columns.

```{r}
library(dplyr)

iris %>% 
  select(Species, everything())
```



---

### Tip 4: Renaming all variables in one go

One command to get them all in lower case. And one more to replace those dots with underscores...


```{r}
library(dplyr)
library(stringr)

iris %>% 
  rename_all(tolower) %>% 
  rename_all(~str_replace_all(., "\\.", "_"))
```

---

### Tip 5: Cleaning up your observations in one go

The `select_all/if/at` and `rename_all/if/at` functions will only modify the variable names, not the observations. If you want to change those, use the `mutate` variant!

```{r}
library(dplyr)
library(stringr)

storms %>% 
  select(name, year, status) %>% 
  mutate_all(tolower) %>% 
  mutate_all(~str_replace_all(., " ", "_"))
```




---

### Tip 6: Finding the 5 highest/lowest values

You can use `top_n` to find the 5 cars with the highest  horsepower without ordering them first.

```{r}
library(dplyr)

mtcars %>% 
  top_n(5, hp)
```

---

### Tip 7: Adding the amount of observations

You can add the amount of observations without summarising them yourself.  
And if you don't like the default column name `n`, just change it again with a `rename()` statement.

```{r}
library(dplyr)

mtcars %>% 
  select(-(drat:vs)) %>% 
  add_count(cyl) %>% rename(n_cyl = n) %>% 
  add_count(am) %>% rename(n_am = n)
```


---

### Tip 8: Making new discrete variables

`case_when()` can be a very powerful tool to make new discrete variables based on other columns.

```{r}
starwars %>%
  select(name, species, homeworld, birth_year, hair_color) %>%
  mutate(new_group = case_when(
      species == "Droid" ~ "Robot",
      homeworld == "Tatooine" & hair_color == "blond" ~ "Blond Tatooinian",
      homeworld == "Tatooine" ~ "Other Tatooinian",
      hair_color == "blond" ~ "Blond non-Tatooinian",
      TRUE ~ "Other Human"))
```

---

### Tip 9: going rowwise...

Mutating with aggregate functions by default will take the average/sum/... of the entire column. Via adding `rowwise()` you can aggregate within an observatioin.

```{r}
iris %>% 
  select(contains("Length")) %>% 
  rowwise() %>% 
  mutate(avg_length = mean(c(Petal.Length, Sepal.Length)))
```

---

### Tip 10: Changing your column names after summarise_if

If you've used the `summarise_all`, `summarise_if` and `summarise_at` variants before, you know that the variable name by default does not get changed.  
If you do what a modified name, you can wrap your function inside `funs()` and add a tag that will be added to the variable name.

```{r}
iris %>% 
  summarise_if(is.numeric, funs(avg = mean))
```

