---
title: "5 forcats tricks"
subtitle: "for @WeAreRLadies"
author: "<br><br>Suzan Baert"
output:
  xaringan::moon_reader:
    css: ["default", "rladies-fonts", "rladies"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: ""
      ratio: '16:9'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

library(dplyr)
```



### Tip 1: Lumping levels together in Other

You can use the `fct_lump()` function from `forcats` to keep the 3 most frequent levels and get everything else coded as 'Other'.


.pull-left[
### Current version

```{r}
starwars %>% 
  count(eye_color)
```
]


.pull-right[
### Using fct_lump version

```{r}
library(forcats)

starwars %>% 
  mutate(eye_color = fct_lump(eye_color, n=3)) %>% 
  count(eye_color)

```
]


---

### Tip 1b: Lumping levels together in Other - but under your control

You can use the `fct_other()` function to control yourself what to keep and what to drop inside "other"


.pull-left[
### Control what to keep separate

```{r}
to_keep <- c("brown", "blue", "hazel")

starwars %>% 
  mutate(eye_color = fct_other(eye_color, keep = to_keep)) %>% 
  count(eye_color)

```

]



.pull-right[
### Control what to drop into "other"

```{r}
to_other <- c("white", "pink", "gold", "unknown", "blue-gray", "dark", "green, yellow", "red, blue")

starwars %>% 
  mutate(eye_color = fct_other(eye_color, drop = to_other)) %>% 
  count(eye_color)
```

]

---