---
title: "3-ish forcats tricks"
subtitle: "for @WeAreRLadies"
author: "<br><br>Suzan Baert"
output:
  xaringan::moon_reader:
    css: ["default", "rladies-fonts", "rladies"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: ""
      ratio: '16:9'
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

library(dplyr)
library(ggplot2)
library(forcats)
```



### Tip 1: Lumping levels together in Other

You can use the `fct_lump()` function from `forcats` to keep the 3 most frequent levels and get everything else coded as 'Other'.


.pull-left[
### Current version

```{r}
starwars %>% 
  count(eye_color)
```
]


.pull-right[
### Using fct_lump version

```{r}
library(forcats)

starwars %>% 
  mutate(eye_color = fct_lump(eye_color, n=3)) %>% 
  count(eye_color)

```
]


---

### Tip 1b: Lumping levels together in Other - but under your control

You can use the `fct_other()` function to control yourself what to keep and what to drop inside "other"


.pull-left[
### Control what to keep separate

```{r}
to_keep <- c("brown", "blue", "hazel")

starwars %>% 
  mutate(eye_color = fct_other(eye_color, 
                               keep = to_keep)) %>% 
  count(eye_color)

```

]



.pull-right[
### Control what to drop into "other"

```{r}
to_other <- c("white", "pink", "gold", "unknown", "blue-gray", "dark", "green, yellow", "red, blue")

starwars %>% 
  mutate(eye_color = fct_other(eye_color, 
                               drop = to_other)) %>% 
  count(eye_color)
```

]

---

### Tip 2: Moving elements of a factor around

```{r include = FALSE}

weekdays <- c("monday", "tuesday", "wednesday", "thursday", "friday", "saturday", "sunday")
weekdays_factor <- ordered(weekdays, levels=weekdays)

```


You can use the `fct_relevel()` to change the order of some elements. 

+ The original vector:

```{r}
weekdays_factor
```

<br><br>

+ To change the starting order, add the elements you want to come first in fct_relevel.   
Note: What does not get mentioned will be added at the end in the current order.

```{r}
fct_relevel(weekdays_factor, "sunday")
```


---


### Tip 2b: Moving elements of a factor around


+ To move elements to other positions add the extra "after" element:

```{r}
fct_relevel(weekdays_factor, "sunday", after = 2)
```

<br><br>

+ To move something to the very end:

```{r}
fct_relevel(weekdays_factor, "sunday", after = Inf)
```


---

### Tip 3: Changing factor levels based on count

Using `fct_infreq()` to change the factor levels based on the frequency it occurs. 


```{r include = FALSE}
weekdays_df <- data.frame(day = sample(weekdays, 100, replace = TRUE))
```


.pull-left[

### No ordering

```{r fig.height=3.5}
ggplot(weekdays_df) + 
  geom_bar(aes(day), fill = "deepskyblue4")
```

]


.pull-right[

### After fct_infreq

```{r fig.height=3.5}
weekdays_df %>% 
  mutate(day = fct_infreq(day)) %>% 
  ggplot() + 
  geom_bar(aes(day), fill = "deepskyblue4")
```

]




